---
title: "cegwas2 report for TRAIT_NAME_HOLDER"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float:
      collapsed: true
      smooth_scroll: true
---

<br>

## Usage

* All strain names were converted to the corresponding isotype name, which can be looked up here: https://www.elegansvariation.org/strain/strain_data.tsv
* This R Markdown file imports data tables within Analysis_Results-Date folder and will knit without any modification.
<br>
* It generates plots (almost) identical to the plots in Analysis_Results-Date folder.
<br>
* Features for the interactive plots: 
    * Place mouse on plot region to show related underlying data.
    * Hold and drag the mouse to select a region to zoom in, double click to reset zoom.
    * Click on legend to show/hide different data groups.
<br>
* ggsave code is kept for users to make customized plots. 
<br>
<br>
<br>


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```



```{r include=FALSE}
library.path <- .libPaths()
library("tidyverse", lib.loc = library.path)
library("rrBLUP", lib.loc = library.path)
library("plotly", lib.loc = library.path)
library("DT", lib.loc = library.path)
```



```{r}
# load trait name
# in nextflow use sed to edit this field and make a copy of the .rmd for each trait
trait_name <- "TRAIT_NAME_HOLDER"   

# load independent tests result
total_independent_tests <- read.table("Genotype_Matrix/total_independent_tests.txt", quote="\"", comment.char="", stringsAsFactors=FALSE)

independent_tests <- total_independent_tests[[1]]

independent_test_cutoff <- -log10(0.05/independent_tests)

```


## Manhatton plot
* #### Every dot is a SNV marker.

* #### Red/pink dots are significant QTL loci based on defined thresholds. 
    * Default BF cutoff -log10(p) > 6. If user supplied threshold exists, it will replace BF.
    * Eigen cutoff: -log10(0.05/independent_tests) where "independent_tests" is calculated from Eigen Decomposition of SNV correlation matrix, stored in Analysis_Results-date/⁩Genotype_Matrix/total_independent_tests.txt

* #### Light blue regions are QTL intervals that contain markers significantly associate with the trait. These are the regions that will be shown in the phenotype-genotype, fine mapping and gene plots below. 


```{r, fig.height=3.5, fig.width=12, out.height=350, out.width=1000}
# load processed mapping data. 
# Note that readr::read_delim will throw parsing errors and put NA in rows that contain actual values in strain/value/allele/var.exp so did not use it
processed_mapping <- read.delim(paste0("Mappings/Data/",trait_name,"_processed_mapping.tsv"), stringsAsFactors=FALSE)


# each trait has a separate processed_mapping file now. So the plotting function and loop is removed
# but do check there is only 1 trait and if not, issue warning:
num_traits = length(unique(dplyr::select(processed_mapping,trait)))

if(num_traits > 1){
    print("WARNING: More than 1 trait in processed_mapping table. Only the first one will be plotted.")
}



# Generate _manplot from Run_Mappings.R

manhattan_plot <- processed_mapping %>%
    dplyr::filter(trait == unique(processed_mapping$trait)[1] & CHROM != "MtDNA") %>%
    dplyr::distinct(marker, .keep_all = T) %>%
    dplyr::mutate(EIGEN_CUTOFF = independent_test_cutoff) %>%
    dplyr::mutate(EIGEN_SIG = ifelse(log10p > BF, "1", 
                                     ifelse(log10p > EIGEN_CUTOFF, "2", "0")) )%>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = POS/1e6, y = log10p) +
    ggplot2::scale_color_manual(values = c("0" = "black", 
                                           "1" = "red",
                                           "2" = "hotpink3")) +
    ggplot2::geom_rect(ggplot2::aes(xmin = startPOS/1e6,    # this is the plot boundary for LD and gene plots
                                    xmax = endPOS/1e6,    # this is the plot boundary for LD and gene plots
                                    ymin = 0, 
                                    ymax = Inf, 
                                    fill = "blue"), 
                       color = "blue",fill = "cyan",linetype = 2, 
                       alpha=.3)+
    ggplot2::geom_hline(ggplot2::aes(yintercept = BF),
                        color = "gray", 
                        alpha = .75,  
                        size = 1) +
    ggplot2::geom_hline(ggplot2::aes(yintercept = EIGEN_CUTOFF),
                        color = "gray", 
                        alpha = .75,  
                        size = 1,
                        linetype = 2) +
    ggplot2::geom_point( ggplot2::aes(color= factor(EIGEN_SIG)) ) +
    ggplot2::facet_grid( . ~ CHROM, scales = "free_x" , space = "free_x") +
    ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(size = 16),
                   axis.text.y = ggplot2::element_text(size = 16),
                   axis.title.x = ggplot2::element_text(size = 20, face = "bold", color = "black", vjust = -0.3), 
                   axis.title.y = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   strip.text.x = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   strip.text.y = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   plot.title = ggplot2::element_text(size = 24, face = "bold", vjust = 1), 
                   panel.background = ggplot2::element_rect(color = "black",size = 1.2),
                   legend.position = "none",
                   strip.background = element_blank()) +
    ggplot2::labs(x = "Genomic Position (Mb)",
                  y = expression(-log[10](italic(p))))


# save the plot to file. keep this code for user to create plots with their own styles
# ggsave(manhattan_plot, 
#        filename = paste0(trait_name, "_manplot.pdf"),
#        height = 4, 
#        width = 12)

# show plot in html output
manhattan_plot
```



```{r}
# read in QTL regions for all traits. Note that QTL_peaks.tsv is generated even if there is no significant regions.
# based on number of regions & width of regions to decide whether generate pxg and fine mapping plots

if (file.info("Mappings/Data/QTL_peaks.tsv")$size!=0) {

    whether_print_peaks <- TRUE
        
    QTL_peaks <- read.delim("Mappings/Data/QTL_peaks.tsv", header=FALSE, stringsAsFactors=FALSE) %>% 
        dplyr::filter(V1==trait_name)  %>% 
        dplyr::rename(Trait=V1, Chrom=V2, Region_start=V3, Peak=V4, Region_end=V5) 
    
    
    if (nrow(QTL_peaks) < 2) { 
        whether_plot_LD <- FALSE 
        } else { whether_plot_LD <- TRUE }
    
    
    if (nrow(QTL_peaks)==0) {
        print("No significant QTL region for this trait")
        whether_fine_map <- FALSE # skip related chunks
        whether_plot_LD <- FALSE
    } else {


        QTL_peaks <- dplyr::mutate(QTL_peaks, Region_width=Region_end - Region_start) %>% dplyr::filter(Region_width < 1500000)
        
        if (nrow(QTL_peaks)==0) {
            print("Significant QTL regions are too wide for plotting")
            whether_fine_map <- FALSE # skip related chunks
            
        } else {
            whether_fine_map <- TRUE # run related chunks
            
        }
    }
    
} else { 
    print("No significant peak for this trait")
    whether_print_peaks <- FALSE
    whether_plot_LD <- FALSE
    whether_fine_map <- FALSE
}
```


```{r, eval=whether_print_peaks}
# print out the table to html
# this doesn't print within if{}
# QTL_peaks was filtered in previous chunk, so reload
QTL_peaks_original <- read.delim("Mappings/Data/QTL_peaks.tsv", header=FALSE, stringsAsFactors=FALSE) %>% 
        dplyr::filter(V1==trait_name)  %>% 
        dplyr::rename(Trait=V1, Chrom=V2, Region_start=V3, Peak=V4, Region_end=V5) 

DT::datatable(QTL_peaks_original)

```

<br>
<br>

## Pair-wish linkage between QTL regions by R square
```{r, fig.width=6, fig.height=4, eval=whether_plot_LD}
ldcalc <- read.table(file = paste0("Mappings/Data/", trait_name, "_LD_between_QTL_regions.tsv"), header = T) %>% as.matrix()

# convert to long format for ggplot
LDs <- tbl_df(data.frame(ldcalc) %>% dplyr::add_rownames(var = "SNP1")) %>%
    tidyr::gather(SNP2, r2, -SNP1) %>% dplyr::arrange(SNP1) %>%
    tidyr::separate(SNP1, sep = "_", into = c("CHROM1","POS1"), remove = F) %>%
    dplyr::arrange(CHROM1, as.numeric(POS1))

ldplot <- ggplot2::ggplot(LDs) +
    ggplot2::aes(x = factor(SNP1, levels = unique(SNP1), ordered = T), y = factor(SNP2, levels = unique(SNP1), ordered = T)) +
    ggplot2::geom_tile(ggplot2::aes(fill = r2)) +
    ggplot2::geom_text(ggplot2::aes(label = signif(r2, 3)), fontface = "bold", size = 4.5) +
    ggplot2::theme(text = ggplot2::element_text(size = 9.5, color = "black"),
                   axis.text = ggplot2::element_text(size = 9.5, color = "black"),
                   axis.text.y = ggplot2::element_text(size = 9.5, color = "black"),
                   axis.title.x = ggplot2::element_blank(),
                   axis.title.y = ggplot2::element_blank(),
                   legend.position = "none") +
    scale_x_discrete(labels = function(x) { gsub("_", ":", x) }, expand = c(0, 0)) + 
    scale_y_discrete(position = "right", labels = function(x) { gsub("_", ":", x) }, expand = c(0, 0)) + scale_fill_continuous(high = "indianred", low = "white", na.value = "white")

ldplot
```

```{r, eval=!whether_plot_LD}
print("Not enough QTL regions to calculate linkage")
```


<br>
<br>

## Divergent region, haplotype and fine mapping for each significant QTL region 
Only for regions less than 1.5Mb in size

```{r, results='asis', eval=whether_fine_map}
link_list <- list()

for (i in seq_len(nrow(QTL_peaks))) { 
    
    region_name <- paste0(QTL_peaks[i,"Chrom"], ".", QTL_peaks[i,"Region_start"], "-", QTL_peaks[i,"Region_end"])
    
    link_list[i] <- paste0("[", region_name, "](cegwas2_report_TRAIT_NAME_HOLDER_region_", region_name, ".html)")
}

for (l in link_list) {
    cat("\n##", l, "\n")
    cat("  \n")
}
```

```{r, eval=!whether_fine_map}
print("No significant QTL regions to plot")
```

<br>
<br>

## Burden mapping plots

#### Every dot represents a gene, red color indicates significance based on corresponding test.
<br>
<br>

+ ### SKAT
```{r, fig.height=3.5, fig.width=10, out.width=900}
# plots from plot_burden.R
# Plot SKAT
skat <- read.delim(paste0("BURDEN/SKAT/Data/", trait_name, ".Skat.assoc"), stringsAsFactors=FALSE) %>%
    dplyr::rowwise() %>%
    dplyr::mutate(CHROM = strsplit(strsplit(RANGE,split = ",")[[1]][1],split = ":")[[1]][1],
                  POS = as.numeric(strsplit(strsplit(strsplit(RANGE,split = ",")[[1]][1],split = ":")[[1]][2],split = "-")[[1]][1]),
                  endPOS = as.numeric(strsplit(strsplit(strsplit(RANGE,split = ",")[[1]][1],split = ":")[[1]][2],split = "-")[[1]][2]))%>%
    dplyr::mutate(size = abs(POS-endPOS)) %>%
    dplyr::filter(CHROM!="MtDNA")%>%
    dplyr::ungroup()%>%
    dplyr::filter(NumVar > 1)%>%
    dplyr::mutate(significant = ifelse(Pvalue < .05/n(), TRUE,FALSE ))%>%
    dplyr::arrange(Pvalue)

skat_plot <- skat%>%
    ggplot()+
    aes(x = POS/1e6, y = -log10(Pvalue), alpha = 0.5, color = significant, plotly_label_geneID = Gene )+
    geom_point(size = 0.8)+
    scale_color_manual(values=c("black","red"))+
    facet_grid(.~CHROM, scales = "free", space = "free")+
    ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(size = 16),
                   axis.text.y = ggplot2::element_text(size = 16),
                   axis.title.x = ggplot2::element_text(size = 20, face = "bold", color = "black", vjust = -0.3), 
                   axis.title.y = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   strip.text.x = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   strip.text.y = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   plot.title = ggplot2::element_text(size = 24, face = "bold", vjust = 1), 
                   panel.background = ggplot2::element_rect(color = "black",size = 1.2),
                   legend.position = "none",
                   strip.background = element_blank())+
    labs(x = "Genomic Position (Mb)", 
         y = "-log[10](p)")

# save plot to file
# ggsave(filename = paste0(trait_name, "_SKAT.pdf"), 
#        plot = skat_plot,
#        height = 4, 
#        width = 12)

# print plot to html output
# print(skat_plot)
ggplotly(skat_plot, tooltip = c("plotly_label_geneID","y")) %>% toWebGL()


# show table in html output. 
# Pvalue is the y-axis in plot
skat_link <- filter(skat, significant == TRUE) %>% 
    mutate(Gene_ID = paste0('<a href=https://wormbase.org/species/c_elegans/gene/', Gene, '>', Gene,'</a>' ) ) %>%
    select(Gene_ID, RANGE, Stat, significant)  

DT::datatable(skat_link, escape = FALSE)
```

<br>
<br>

+ ### Variable Threshold
```{r, fig.height=3.5, fig.width=10, out.width=900}
##############
# PLOT VTprice
benreg_burden <- read.delim(paste0("BURDEN/VT/Data/", trait_name, ".VariableThresholdPrice.assoc"), stringsAsFactors=FALSE)
benreg_burden$Gene <- as.character(benreg_burden$Gene)
benreg_burden$RANGE <- as.character(benreg_burden$RANGE)

benreg_burden<-benreg_burden%>%
    dplyr::rowwise()%>%
    dplyr::mutate(CHROM = strsplit(strsplit(RANGE,split = ",")[[1]][1],split = ":")[[1]][1],
                  POS = as.numeric(strsplit(strsplit(strsplit(RANGE,split = ",")[[1]][1],split = ":")[[1]][2],split = "-")[[1]][1]),
                  endPOS = as.numeric(strsplit(strsplit(strsplit(RANGE,split = ",")[[1]][1],split = ":")[[1]][2],split = "-")[[1]][2]))%>%
    dplyr::mutate(size = abs(POS-endPOS)) %>%
    ungroup()%>%
    dplyr::mutate(significant = ifelse(PermPvalue < .05/n(), TRUE,FALSE ))%>%
    dplyr::filter(CHROM!="MtDNA",NumVar>1,size >500)

burden_manplot <- benreg_burden%>%
    ggplot()+
    aes(x = POS/1e6, y = Stat, size = NumVar, alpha = 0.5, color = significant, plotly_label_geneID = Gene) +
    geom_point(size = 0.8)+
    scale_color_manual(values=c("black","red"))+
    facet_grid(.~CHROM, scales = "free", space = "free")+
    theme_bw()+
    ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(size = 16),
                   axis.text.y = ggplot2::element_text(size = 16),
                   axis.title.x = ggplot2::element_text(size = 20, face = "bold", color = "black", vjust = -0.3), 
                   axis.title.y = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   strip.text.x = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   strip.text.y = ggplot2::element_text(size = 20, face = "bold", color = "black"), 
                   plot.title = ggplot2::element_text(size = 24, face = "bold", vjust = 1), 
                   panel.background = ggplot2::element_rect(color = "black",size = 1.2),
                   legend.position = "none",
                   strip.background = element_blank())+
    labs(x = "Genomic Position (Mb)", y = "Test Statisitic")

# save plot to file
# ggsave(filename = paste0(trait_name, "_VTprice.pdf"), 
#        plot = burden_manplot,
#        height = 4, 
#        width = 12)


# show plot in html output
# print(burden_manplot)
ggplotly(burden_manplot, tooltip = c("plotly_label_geneID","y")) %>% toWebGL()


# show table in html output. 
# Stat is the y-axis in plot. what is it??
benreg_burden_link <- filter(benreg_burden, significant == TRUE) %>% 
    mutate(Gene_ID = paste0('<a href=https://wormbase.org/species/c_elegans/gene/', Gene, '>', Gene,'</a>' ) ) %>%
    select(Gene_ID, RANGE, Stat, significant)  

DT::datatable(benreg_burden_link, escape = FALSE)
```

<br>
<br>

+ ### Genes significant in both statistical tests 
```{r}
skat_sig <- dplyr::filter(skat, significant == TRUE) %>% dplyr::rename(SKAT_stat = Stat)

benreg_burden_sig <- dplyr::filter(benreg_burden, significant == TRUE)  %>% dplyr::rename(VariableThreshold_stat = Stat)

both_sig <- inner_join(skat_sig, benreg_burden_sig, by=c("Gene","RANGE")) %>% 
    mutate(Gene_ID = paste0('<a href=https://wormbase.org/species/c_elegans/gene/', Gene, '>', Gene,'</a>' ) ) %>%
    select(Gene_ID, RANGE, SKAT_stat, VariableThreshold_stat)  

DT::datatable(both_sig, escape = FALSE) 
```

<br>
<br>

## Please kindly cite the following publications
* CeNDR paper
* cegwas2 paper
* divergent region paper
* haplotype paper

<br>
<br>

```{r}
sessionInfo()
```
